---
title: "Placeholder Golf Project"
author: "Thomas Schechter and Ian Bogley"
date: "12/10/2021"
output: pdf_document
---

```{r}
#Package unloading
library(pacman)
p_load(tidyverse,data.table,tidymodels)

#Read in data
pga_data17 <- fread("PGATOUR_data2.csv")
```
Read data in and installed packages.
```{r}
#Subset data
pga_data <- pga_data17[1:195,]

```
Subset the data to remove null and na values, turned subset data into data.table
```{r}
#Begin exploring the relationships in the data

#Correlation b/w rounds played and top tens
model1 <- lm(ROUNDS_PLAYED~NUMBER_OF_TOP_Tens, pga_data)

tidy(model1)

#Plot the previous relationship
ggplot(pga_data, aes(ROUNDS_PLAYED, NUMBER_OF_TOP_Tens)) + 
  geom_point()
```
Not a significant amount of visual correlation between number of rounds played and number of top ten finishes.

```{r}
#Plotting strokes gained putting vs. top ten finishes
putt_plot <- ggplot(pga_data, aes(SG_PUTTING_PER_ROUND, NUMBER_OF_TOP_Tens)) +
  geom_point()

putt_plot
```
Strokes gained putting seems to have more of an effect on success.

```{r}
#Plotting average drive distance vs. top ten finishes
drive_plot <- ggplot(pga_data, aes(AVG_Driving_DISTANCE,NUMBER_OF_TOP_Tens)) + geom_point()

drive_plot
```
Some relationship appears present, though with more outliers on both ends of the spectrum.

```{r}

```

